---
title: "R-club-May-3.Benn"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error =  TRUE)
```

```{r}
library(tidyverse)
```  

##3.6.1 Exercises  

**1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?**
geom_line(), geom_boxplot(), geom_histogram(), geom_area().

**2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.**  

This will produce a scatterplot of displacement by highway mileage, with points colored based on drivetrain. There will also be a trendline with no standard error bars.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```  

So I was not quite correct. The trendlines also follow the color aes, so there are 3 lines.

**3. What does show.legend = FALSE do? What happens if you remove it? Why do you think I used it earlier in the chapter?**  

It hides the legend. This keeps the graph looking cleaner if there are many categories.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point(show.legend = FALSE) + 
  geom_smooth(se = TRUE)
```  

Not clear on this one.

**4. What does the se argument to geom_smooth() do?**
It generates standard errors around the trendlines.

**5. Will these two graphs look different? Why/why not?**
The graphs should look the same - they do the mapping, just in different places.

**6. Generate R code to recreate graphs.**
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```  
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv), se = FALSE)
``` 
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv), se = FALSE)
``` 
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_smooth(mapping = aes(x = displ, y = hwy), se = FALSE)
``` 
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv), se = FALSE)
``` 
```{r}
ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy, color = drv, stroke = 4))
``` 

##3.7.1 Exercises
**1. What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?**
Stat_summary is associated with geom_histogram and geom_freqpoly.

```{r}
ggplot(data=mpg) +
  geom_histogram(mapping = aes(x=hwy),binwidth = 5)
```  

**couldn't figure out how to duplicate the above with stat_summary()**
```{r}
ggplot(data=mpg) +
  stat_summary(fun.y = "count", mapping = aes(hwy))
```  

**couldn't figure out how to duplicate the stat_summary thing with either of the associated geoms**
```{r}
ggplot(data = diamonds) + 
  geom_histogram(mapping = aes(x = cut, y = depth))
```

**2. What does geom_col() do? How is it different to geom_bar()?**
geom_col() doesn't seem to exist in my version of R.

**3. Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?**

geom_boxplot - stat_boxplot
geom_bar - stat_count
geom_point - stat = "identity"
geom_errorbar - stat = "identity"

I'm not sure what the last part of the question is getting at.

**4. What variables does stat_smooth() compute? What parameters control its behaviour?**
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth()
```
smooth() plots a trendline and produces error bars. It takes input in terms of the method - either a formula or a method.

**5. In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?**  

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop..))
```

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
```

It looks like without group, it's calculating proportions within each diamond grade.

##3.8.1 Exercises

**1. What is the problem with this plot? How could you improve it?**
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()
```  

I'm guessing it's omitting points.
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point() +
  geom_jitter()
```  

**2. What parameters to geom_jitter() control the amount of jittering?**
width and height control the amount of jittering.

**3. Compare and contrast geom_jitter() with geom_count().**
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point() +
  geom_count()
```  

jitter shifts points slightly to prevent overplotting, while count increases point size when there are multiple points at a particular coordinate.

**4. What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.**
The default is dodge.
```{r}
ggplot(data = mpg, mapping = aes(x = drv, y = hwy, fill = manufacturer)) + 
  geom_boxplot(position = position_dodge(1))
```  

##3.9.1 Exercises

**1. Turn a stacked bar chart into a pie chart using coord_polar().**
```{r}
x <- ggplot(data = mpg) + geom_bar(mapping = aes(x = manufacturer, fill = drv))
x + coord_flip()
x + coord_polar()
```  

**2. labs()?**
Allows you to specify labels.

**3. coord_quickmap vs coord_map**
Quickmap makes a simplified projection that runs faster, but is less accurate.

**4. What does the plot below tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?**
This tells us that city and hwy mpg are correlated.
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline()
```  

I can't tell the difference when coord_fixed() is removed - the scales stay the same, though the proportions are different. Not sure why this is significant.
I'm also not clear on what abline is doing - it's clearly drawing a line, but I'm not sure what the defaults are for setting the slope and intercept.

##4.4 Practice
**1. Why does this code not work?**
my_variable is mispelled in line 2- there's a 1 instead of an i.

**2. Tweak each of the following R commands so that they run correctly:**

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

```{r}
filter(mpg, cyl == 8)
```

**4. alt+shift+k**
This brings up keyboard shortcuts. This can also be accessed via the help menu.

**graph my own data!**

```{r}
setwd("~/Box Sync/BIS2C/2C Analysis/App evaluation")
practical <- read.delim("F16.lab.practical.txt",header = T)
ggplot(practical, aes(x=MT.Letter,y=Practical.Live.Plants)) + 
  geom_boxplot(aes(fill=App.completion))+
  ggtitle("Performance on Questions Requiring Live Plant ID")+ 
  labs(x="Midterm Letter Grade", y="Live Plant Questions Score")
```
