---
title: "Benn_R_Club_May_10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(nycflights13)
data <- flights
```

##5.2.4 Exercises

**1. Find all flights that**

1. Had an arrival delay of two or more hours  

```{r}
q1 <- filter(data, arr_delay >= 120)
```

2. Flew to Houston (IAH or HOU)  

```{r}
q2 <- filter(data, dest == "IAH" | dest == "HOU")
```

3. Were operated by United, American, or Delta  

```{r}
q3 <- filter(data, carrier == "UA"| carrier == "AA" | carrier == "DL")
q3.2 <- filter(data, carrier %in% c('UA','AA','DL'))
```

4. Departed in summer (July, August, and September)

```{r}
q4 <- filter(data, month %in% c(7,8,9))
```

5. Arrived more than two hours late, but didn’t leave late

```{r}
q5 <- filter(data, arr_delay > 120 & dep_delay == 0)
```

6. Were delayed by at least an hour, but made up over 30 minutes in flight

```{r}
q6 <- filter(data, dep_delay >= 60 & (dep_delay - arr_delay) > 30)
```

7. Departed between midnight and 6am (inclusive)
```{r}
q7 <- filter(data, dep_time <= 600)
```

**2.Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?**

Between looks for numeric values in between two numbers.
```{r}
q4.2 <- filter(data, between(dep_time, 0, 600))
```

**3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?**

```{r}
q3.3 <- filter(data, is.na(dep_time))
```

These data points are also missing arrival times, air times and delays. These are likely canceled flights.

**4. Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)**

```{r}
NA ^ 0
NA | TRUE
NA & FALSE
NA*0
```
Anything raised to the 0 is 1, NA or TRUE would evaluate to true as the first position doesn't matter, the third one I'm less sure of. I can't figure out the general rule.

##5.3.1 Exercises

**1. How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).**

```{r}
q3.1.1 <- arrange(data, desc(is.na(dep_time)))
```

**2. Sort flights to find the most delayed flights. Find the flights that left earliest.**

```{r}
q3.1.2 <- arrange(data, desc(dep_delay))
q3.1.2b <- arrange(data, dep_delay)
```

**3. Sort flights to find the fastest flights.**

```{r}
q3.1.3 <- arrange(data, desc(distance/air_time))
```

**4. Which flights travelled the longest? Which travelled the shortest?**

```{r}
q3.1.4a <- arrange(data, desc(distance))
```

The JFK to Honolulu flights were the longest.

```{r}
q3.1.4b <- arrange(data, distance)
```

The shortest regular flights were between Newark and Philly.

##5.4.1 Exercises

**1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.**

```{r}
select(data, dep_time, dep_delay, arr_time, arr_delay)
select(data, dep_time:arr_delay, -(contains("sched")))
select(data, contains("arr_"), contains("dep_"), -(contains("sched")))
select(data, ends_with("time"), ends_with("delay"),-(contains("sched")), -(contains("air")))
```

**2. What happens if you include the name of a variable multiple times in a select() call?**

```{r}
select(data, dep_time, dep_time, dep_time)
```

It seems to ignore it.

**3. What does the one_of() function do? Why might it be helpful in conjunction with this vector?**

```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(data, one_of(vars))
```

It looks for column names that contain any of the strings in the vector.

**4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?**

```{r}
select(data, contains("TIME"))
```

This ignores case and returns everything with time.

```{r}
select(data, contains("TIME", ignore.case = FALSE))
```

##5.5.2 Exercises

**1. Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.**

```{r}
q5.2.1 <- mutate(data, dep_abs = ((dep_time %/% 100)*60) + dep_time %% 100, sched_dep_abs = ((sched_dep_time %/% 100)*60) + sched_dep_time %% 100)
```

**2. Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?**
```{r}
q5.2.2 <- mutate(data, air_time, air.diff = arr_time - dep_time)
```

I'm not sure how air_time is being calculated - it doesn't seem to line up with the listed times.

**3. Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?**
dep_delay is the difference between dep_time and sched_dep_time.

**4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().**

```{r}
q5.2.4 <- mutate(data, delay.rank = min_rank(desc(dep_delay, arr_delay)))
q5.2.4 <- arrange(q5.2.4, delay.rank)
filter(q5.2.4, delay.rank <= 10)
```

So it looks like you can set the tie method using ties.method, however I can't find a list of all the options anywhere.

**5. What does 1:3 + 1:10 return? Why?**

```{r}
1:3 + 1:10
```

It returns a vector, where 1, 2, and 3 were added to the numbers 1-10, with recycling.

**6. What trig functions does R have?**

Sin, cos, tan, arcsin, arccos, arctan.
