---
title: "Benn.R-club-May-17"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(nycflights13)
data <- flights
```

```{r}
sessionInfo()
```

##5.6.7 Exercises

**1. Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios:**

A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.

A flight is always 10 minutes late.

A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.

99% of the time a flight is on time. 1% of the time it’s 2 hours late.

Which is more important: arrival delay or departure delay?

```{r}
not_cancelled <- flights %>%
  filter(!is.na(dep_delay), !is.na(arr_delay))
```
This question is kind of vague.

**2. Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).**

```{r}
not_cancelled %>% 
  group_by(dest) %>%
  summarise(flights = n())

not_cancelled %>% 
  group_by(tailnum) %>%
  summarise(total = sum(distance))
```  

**3. Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay) ) is slightly suboptimal. Why? Which is the most important column?**
It seems like we would only really care about the arr_delay - i.e. the flights that never made it. So we could dispense with asking about missing dep_delay information.

**4. Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?**

```{r}
flights %>%
  group_by(day) %>%
  summarise(
    flights = n(),
    cancelled = sum(is.na(arr_delay)),
    avg_delay = mean(arr_delay, na.rm = TRUE)
  ) %>%
  arrange(desc(cancelled)) %>%
  ggplot(aes(x = avg_delay, y = cancelled)) + geom_point()
```  

It looks like the number of cancelled flights is positively correlated with average delay. This makes sense - a storm or computer glitch that causes delays will also lead to cancellations.

**5. Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))**

```{r}
flights %>%
  group_by(carrier) %>%
  summarise(avg_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(avg_delay))
```  

Frontier (F9) has the highest average delay.

```{r}
flights %>%
  group_by(carrier, dest) %>%
  summarise(avg_delay = mean(arr_delay, na.rm = TRUE)) %>%
  ggplot(aes(x = dest, y = avg_delay, fill = carrier)) + geom_bar(stat = "identity", position = "dodge")
```
It seems like it's complicated because certain carriers utilizes specific airports more heavily. This could be addressed by using statistical approaches.

**6. What does the sort argument to count() do. When might you use it?**

It sorts the results of count from highest to lowest. I would probably use it as a default.

##5.7.1 Exercises

**1. Refer back to the lists of useful mutate and filtering functions. Describe how each operation changes when you combine it with grouping.**

**2. Which plane (tailnum) has the worst on-time record?**

```{r}
flights %>%
  group_by(tailnum) %>%
  summarise(mean_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(mean_delay))
```

N844MH has the highest average delay.

```{r}
flights %>%
  group_by(tailnum) %>%
  filter(arr_delay > 0) %>%
  count(tailnum) %>%
  arrange(desc(n))
```

N725MQ has the most delayed flights.

```{r}
flights %>%
  group_by(tailnum) %>%
  filter(arr_delay > 60) %>%
  count(tailnum) %>%
  arrange(desc(n))
```

N10575 has the most flights delayed an hour or more.

**3. What time of day should you fly if you want to avoid delays as much as possible?**
```{r}
data %>%
  group_by(hour) %>%
  summarise(mean_delay = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(mean_delay)
```

It looks like late night flights are most likely to be delayed, while early morning flights tend to leave slightly early.

**4. For each destination, compute the total minutes of delay. For each, flight, compute the proportion of the total delay for its destination.**

```{r}
not_cancelled %>%
  group_by(dest) %>%
  summarise(total_delay = sum(arr_delay)) %>%
  arrange(desc(total_delay))
```
```{r}
not_cancelled %>%
  group_by(dest) %>%
  mutate(prop_delay = arr_delay / sum(arr_delay))
```

**5. Delays are typically temporally correlated: even once the problem that caused the initial delay has been resolved, later flights are delayed to allow earlier flights to leave. Using lag() explore how the delay of a flight is related to the delay of the immediately preceding flight.**

```{r}
not_cancelled %>%
  group_by(origin) %>%
  mutate(lag = lag(dep_delay, na.rm = TRUE)) %>%
  select(flight, origin, dep_delay, lag) %>%
           ggplot(aes(x=dep_delay, y=lag)) + geom_point()
```

