---
title: "Week 3.Benn"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

##3.2.4 Exercises
1. Run ggplot(data = mpg) what do you see?  
```{r}
ggplot(data = mpg)
```
No plot is produced.  

2. How many rows are in mtcars? How many columns?
```{r}
nrow(mtcars)
```
```{r}
ncol(mtcars)
```

3.What does the drv variable describe? Read the help for ?mpg to find out.  
drv describes whether the car is front, rear, or 4-wheel drive.  

4. Make a scatterplot of hwy vs cyl.  
```{r}
ggplot(data=mpg) + geom_point(mapping = aes(x = cyl, y = hwy))
```


5. What happens if you make a scatterplot of class vs drv. Why is the plot not useful?
```{r}
ggplot(data=mpg) + geom_point(mapping = aes(x = drv, y = class))
```

The graph basically shows that most drive types are represented in most classes.  

##3.3.1 Exercises  
1. What’s gone wrong with this code? Why are the points not blue?  
The color assignment is within the aes(), so the code is trying to assign color based on a non-existant "blue " variable.  

2. Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?  

Manufacturer, model, cyl, trans, drv, fl, and class are categorical. Displ, cty, and hwy are continuous. Some information is provided via the type of atomic vector - character vectors are certainly categorical, however integers may be categorical or continuous.  

3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?  

```{r}
ggplot(data=mpg) + geom_point(mapping = aes(x = cyl, y = hwy, color = displ))
```  
```{r}
ggplot(data=mpg) + geom_point(mapping = aes(x = cyl, y = hwy, size = displ))
```  
```{r}
ggplot(data=mpg) + geom_point(mapping = aes(x = cyl, y = hwy, shape = displ))
```  

4. What happens if you map the same variable to multiple aesthetics?
```{r}
ggplot(data=mpg) + geom_point(mapping = aes(x = cyl, y = hwy, color = displ, color = displ))
```  

5. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)